Technical Report

### 2017-10-25 fully-abstract-comp/technical-report.tex
a few additions - the LR is still a mess.
updates to step relation to take special case into consideration
A3
Propose a new magic-wand-style quantifications in the expression relation.
made several small changes based on meeting with Dominique

\documentclass[a4paper]{article}
\usepackage[a4paper]{geometry}
\usepackage[utf8]{inputenc}

@@

%%% Operational semantics
\newcommand{\step}{\rightarrow}
\newcommand{\nstep}[1][n]{\step_{#1}}
\newcommand{\steps}{\step^*}
\newcommand{\diverge}{\Uparrow}
\newcommand{\term}[1][-]{\Downarrow}

@@

%%% Permissions
\newcommand{\plainperm}[1]{\textsc{#1}}
%\newcommand{\rwlx}{\plainperm{rwlx}}
\newcommand{\rwx}{\plainperm{rwx}}
\newcommand{\rx}{\plainperm{rx}}
%\newcommand{\rwlxo}{\plainperm{rwlxo}}
%\newcommand{\rwlo}{\plainperm{rwlo}}
%\newcommand{\rwl}{\plainperm{rwl}}
%\newcommand{\rwxo}{\plainperm{rwxo}}
%\newcommand{\rwo}{\plainperm{rwo}}
\newcommand{\rw}{\plainperm{rw}}
%\newcommand{\rxo}{\plainperm{rxo}}
\newcommand{\readonly}{\plainperm{r}}
\newcommand{\ro}{\readonly}
\newcommand{\noperm}{\plainperm{0}}
%\newcommand{\nopermo}{\plainperm{0o}}
\newcommand{\enter}{\plainperm{e}}
%\newcommand{\entero}{\plainperm{eo}}

@@

\newcommand{\encType}[1]{\plainfun{encodeType}{#1}}
\newcommand{\encPerm}[1]{\plainfun{encocePerm}{#1}}
\newcommand{\decPerm}[1]{\plainfun{decodePerm}{#1}}
\newcommand{\encInstr}[1]{\plainfun{encodeInstruction}{#1}}
\newcommand{\decInstr}[1]{\plainfun{decodeInstruction}{#1}}
\newcommand{\updPcAddr}[1]{\plainfun{updatePc}{#1}}
\newcommand{\updPcPerm}[1]{\plainfun{updatePcPerm}{#1}}
\newcommand{\nonExec}[1]{\plainfun{nonExecutable}{#1}}
\newcommand{\exec}[1]{\plainfun{executable}{#1}}
%\newcommand{\opaquePerm}[1]{\plainfun{opaquePerm}{#1}}
\newcommand{\readAllowed}[1]{\plainfun{readAllowed}{#1}}
\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}{#1}}
%\newcommand{\writeLocalAllowed}[1]{\plainfun{writeLocalAllowed}{#1}}
%\newcommand{\nonLoc}[1]{\plainfun{nonLocal}{#1}}
\newcommand{\nonLinear}[1]{\plainfun{nonLinear}{#1}}
\newcommand{\linCons}[1]{\plainfun{linearityConstraint}{#1}}
%\newcommand{\isLoc}[1]{\plainfun{isLocal}{#1}}
\newcommand{\isLinear}[1]{\plainfun{isLinear}{#1}}
\newcommand{\sealAss}[1]{\plainfun{sealAssignment}{#1}}
\newcommand{\noRetStkReg}[1]{\plainfun{noRetStk_{reg}}{#1}}
\newcommand{\noRetStkMs}[1]{\plainfun{noRetStk_{ms}}{#1}}
\newcommand{\withinBounds}[1]{\plainfun{withinBounds}{#1}}
\newcommand{\nonZero}[1]{\plainfun{nonZero}{#1}}
\newcommand{\callCond}[1]{\plainfun{callCondition}{#1}}

@@

% \dominique{7-9-2017: Should we keep local/global and linear/non-linear?  I was thinking just the latter would suffice...}
% \lau{07-09-2017: The local/global mechanism still has some value as it now can be used for the purpose of arguments that cannot be stored by the receiver (if you are willing to ``pay'' the clearing cost). I do, however, not think that it is strictly necessary here to make things work.}
% \lau{12-09-2017: I will remove the local capabilities today for the reasons we have discussed (they add to the complexity and they seem to be of little to no use without a clear memory instruction).}
Source language domains:
\[
  \begin{array}{rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
    &\Word & \defeq & \ints \uplus \SealableCaps\\
    \permbnf \in& \Perm & \defbnf & \dots \\
    &\linbnf & \defbnf & \linear \mid \normal \\
    &\aendbnf & \in & \Addr \uplus \{\infty \} \\
    &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
    &\scbnf &\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\linbnf,\sealbasebnf,\sealendbnf,\sigma}\\
    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd(\sigma) \mid \retptrc(\sigma)}\\ 
    &\rbnf & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
    &\Reg & \defeq & \RegName \fun \Word\\
    &\Mem & \defeq & \Addr \fun \Word \\
    &\MemSeg & \defeq & \Addr \parfun \Word \\
    {\sourcecolor \sstkframe \in} & {\sourcecolor \StkFrame} & {\sourcecolor \defeq} & {\sourcecolor \Caps \times \MemSeg \times \Addr}\\
    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus (\{\halted\} \times \Mem)
  \end{array}
\]
\lau{06-09-2017: Should seals also have the option to be linear?}
\lau{13-09-2017: For now I have added linearity to the seals.}
\dominique{27-9-2017: I think we should have versions of $\retptrd$ and $\retptrc$ for non-top stack frames and define that invoking them with the current stack pointer will fail.}
\lau{28-09-2017: Why would you like this? If we want to add this, we will have to be careful, because an old code-return-pointer may fit perfectly well with our new data-return-pointer, so in this case it should not fail. We added the seal to the return pointers, so we had some way of figuring out whether two return pointers would actually fit together.}
\dominique{27-9-2017: Likewise, I think we need to have a source-language representation for a pointer into one of the non-top stack frames, as we may need to be able to back-translate adversary code that reads/writes in those.}
\lau{28-09-2017: At the moment, the only part of the stack that is stored in the stack frame is the private part. Specifically it is the part from the base to the current address of the stack pointer \emph{in $r_\stk$}. This means that if someone saves a stack pointer, so it is not included in the call, then they can still use it because that part of the stack was not encapsulated in a stack frame. Makes sense?}
The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
\begin{itemize}
\item $\linbnf$ defines domain $\Linear$
\item $\scbnf$ defines domain $\SealableCaps$
\item $\cbnf$ defines domain $\Caps$
\item $\rbnf$ defines the finite set $\RegName$.
\item $\Perm$ is defined as the set of permissions in Figure~\ref{fig:perm-hier}.
\end{itemize}

@@

\subsection{Permissions}
\lau{23-03-2017: I am not sure we need the opaque capabilities anymore.} % I am not sure why I did not think the opaque capabilities where necessary, but I have a feeling we talked about this in Leuven and came to the conclusion that they were necessary.
\begin{figure}[!h]
  \centering
  \begin{tikzpicture}[main node/.style={}]
    \node[main node] (rwx) {$\rwx$};
    \node[main node] (rx) [below of=rwx] {$\rx$};
    \node[main node] (e) [below of=rx] {$\enter$};

    \node[main node] (rw) [below left of=rwx,xshift=-0.5cm,yshift=-0.3cm] {$\rw$};
    \node[main node] (r) [below of=rw] {$\readonly$};
    \node[main node] (0) [below of=r] {$\noperm$};

@@

    (rwx) edge (rx)
    (rx) edge (e)

    (rw) edge (rwx)
    (r) edge (rx)
    (0) edge (e);
  \end{tikzpicture}

@@

\begin{align*}
  \src{\Phi} & \; \src{\step \sem{\scall{r_1}{r_2}}(\Phi)} &  &\sourcecolor\left.\arraycolsep=0pt
                                                  \begin{array}[t]{l}
                                                    \text{if }\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr) \tand \\
                                                    \callCond{\Phi,r_1,r_2,\aaddr} \tand\\
                                                    \baddr \leq \aaddr \tand \aaddr + \calllen-1 \leq \eaddr \tand \\
                                                    \exec{\Phi(\pcreg)}
                                                  \end{array}\right.\\
  \Phi & \step \sem{\decInstr{\Phi.\mem(\aaddr)}}(\Phi) & &\left.\arraycolsep=0pt
                                                  \begin{array}[t]{l}
                                                    \text{if }\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr) \tand \\
                                                    \neg\callCond{\Phi,r_1,r_2,\aaddr} \tand\\
                                                    \baddr \leq \aaddr \leq \eaddr\tand\\
                                                    \exec{\Phi(\pcreg)}
                                                  \end{array}\right.\\
  \Phi& \step \failed & & \totherwise
\end{align*}

@@

\[
  \updPcPerm{c} \defeq
  \begin{cases}
    ((\rx,\lin),\baddr,\eaddr,\aaddr) & c = ((\enter,\lin),\baddr,\eaddr,\aaddr)\\
    c & \totherwise
  \end{cases}
\]

@@

\noindent\textbf{jmp and jnz}\\
\begin{align*}
  \sem{\tjmp{r}}(\Phi) = &  
                     \begin{cases}
                       \Phi\updReg{r}{w}\updReg{\pcreg}{\updPcPerm{\Phi(r)}} & w = \linCons{\Phi(r)}
                     \end{cases}
\end{align*}

@@

\noindent\textbf{cca}\\
\emph{Change Current Address}
\lau{15-09-2017: This is the old lea. I changed the name, so it actually reflects what the instruction does. We might also want to add a real lea at some point.}
\begin{align*}
  \sem{\tcca{r_1}{r_2}}(\Phi) = & 
                                  \begin{cases}
                                    \updPcAddr{\Phi\updReg{r_1}{c}} &  
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$}\\
                                      \quad\text{ and in either case $n \in \ints$ and} \\
                                      \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                      \perm \neq \enter \tand \\
                                      c = ((\perm,\lin),\baddr,\eaddr,\aaddr + n)
                                    \end{array}
                                    \\
                                    \updPcAddr{\Phi\updReg{r_1}{c}} &  
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$}\\
                                      \quad\text{ and in either case $n \in \ints$ and} \\
                                      \Phi(r_1) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma} \tand \\
                                      c = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma + n}
                                    \end{array}
                                    \\
                                    \src{\updPcAddr{\Phi\updReg{r_1}{c}}} &  
                                    \sourcecolor\left.
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \text{either $n = r_2$ or $n = \Phi(r_2)$}\\
                                      \quad\text{ and in either case $n \in \ints$ and} \\
                                      \Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
                                      \perm \neq \enter \tand \\
                                      c = \stkptr{\perm,\baddr,\eaddr,\aaddr + n}
                                    \end{array}\right.
                                    \\
                                    \failed & \totherwise
                                  \end{cases}
\end{align*}

@@

\noindent\textbf{seta2b}\\
\begin{align*}
  \sem{\tsetatob{r_1}{r_2}}(\Phi) = & 
                                \begin{cases}
                                  \updPcAddr{}\left(
                                    \arraycolsep=0pt
                                    \begin{array}{rl}
                                    \Phi&\updReg{r_2}{w_2}\\
                                        &\updReg{r_1}{c}
                                    \end{array} \right)
&
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\_) \tand \\
                                      \perm \neq \enter \tand \\
                                      w_2 = \linCons{\Phi(r_2)}\\
                                      c = ((\perm,\lin),\baddr,\eaddr,\baddr)
                                    \end{array} \\
                                  \updPcAddr{}\left(
                                    \arraycolsep=0pt
                                    \begin{array}{rl}
                                    \Phi&\updReg{r_2}{w_2}\\
                                        &\updReg{r_1}{c}
                                    \end{array} \right)
&
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \Phi(r_2) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\_} \tand \\
                                      w_2 = \linCons{\Phi(r_2)}\\
                                      c = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}
                                    \end{array} \\
                                    \failed & \totherwise
                                \end{cases}
\end{align*}

@@

                                   \perm \neq \enter \tand \\

@@

                                   \perm \neq \enter \\

@@

\section{Logical Relation}
\lau{19-10-2017: Rest of source configuration needs to come from somewhere in $\lre$.}
\lau{19-10-2017: Maybe linearity should be taken care of in just one place (instead of in $\lrr$ relation and mem staisfaction). That is, add a predicate that checks whether the configuration satisfies linearity (such a condition may be to syntactic to work?).}
In the following definitions, \src{blue} is used to indicate values related to the source machine. This is unlike previous definitions, where \src{blue} is used  indicate source language specific parts of definitions.
\[
  \lre(W) = \left\{ \npair{\stpair{v}{v}} \middle| 
    \begin{array}{l}
      \forall n' \leq n, \src{k_{d,S}}, \src{k_{c,S}}, k_{d,T}, k_{c,T}, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
      \quad \exists W_\lrp, W_\lrk, W_\lrr, W_\lrm \text{that is a separarion of $W$}\ldotp  \wedge \\
      \qquad\npair[n']{\stpair{v}{v}} \in \lrp(W_\lrp) \wedge \\
      \qquad\npair[n']{\stpair[.]{(k_{d,S},k_{c,S})}{(k_{d,T},k_{c,T})}} \in \lrk(W_\lrk) \wedge \\
      \qquad\npair[n']{\stpair[]{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
      \qquad\Rightarrow\\
      \qquad\npair[n']{\left(\array{l}\src{(\reg_S\updReg{\pcreg,\rretc,\rretd}{v_S,k_{c,S},k_{d,S}},\ms_S,\stk, \ms_\stk)},\\
                                          (\reg_T\updReg{\pcreg,\rretc,\rretd}{v_T,k_{c,T},k_{d,T}},\ms_T)\endarray\right)}
      \in \lro(W)
    \end{array}
            \right\}
\]
\lau{24-10-2017: Dominique, I do not see how the splitting of worlds is going to work out. I can see how it will enforce linearity, but is you give a "pure" region (the kind of region that is only allowed to go to one part of a split) to the $\lrr$-relation, then the memory satisfation does not have it, and it cannot find a memory segment for this part. In other words, the memory satisfaction will not be able to find a concrete memory segment for segments governed by a linear capability in the register file. Maybe the memory satisfaction should have all of the parts, so it can represent them (but also limit what linear capabilities can be used).}

\[
  \lro(W) = \left\{ \npair{\stpair[.]{(\reg_S,\ms_S,\stk_S,\ms_{\stk,S})}{\reg_R,\ms_R}} \middle|
    \begin{array}{l}
      \forall i \leq n, \stpair{\mem}{\mem} \ldotp \\
      \quad \src{(\reg_S,\ms_S \cup \mem_S,\stk_S,\ms_{\stk,S})} \term[i] \Rightarrow (\reg_R,\ms_R) \term\\
      \quad \tand\\
      \quad (\reg_R,\ms_R) \term[i] \Rightarrow \src{(\reg_S,\ms_S \cup \mem_S,\stk_S,\ms_{\stk,S})} \term
    \end{array}
\right\}
\]

\lau{19-10-2017: $\lrk$ may indicate that $\lre$ should really be a relation between a \src{source pc} and a pair of a target pc and stack pointer? Otherwise, what happens with the data part of the return pointer?}
\[
  \lrk(W) = \left\{ \npair{\stpair[.]{\retptrd{\sigma},\retptrc{\sigma}}{k_{d,T},k_{c,T}}} \middle|
    \begin{array}{l}
      k_{d,T} \text{ is a stack pointer with correct end and such with everything related to the source stack} \\
      k_{c,T} \text{ relates to the old pc on the stack}\\
      \forall n' \leq n \ldotp \\
      \quad \forall W' \privft W \ldotp \\
      \qquad \dots \in \lre(W')
    \end{array}
            \right\}
\]

\[
  \lrp(W) = \left\{
    \npair{\stpair{v}{v}} \middle| \dots
  \right\}
\]

>>>>

\documentclass[a3paper]{article}
\usepackage[a3paper]{geometry}
\usepackage[utf8]{inputenc}

@@

%%% Operational semantics
\newcommand{\step}{\rightarrow}
\newcommand{\nstep}[1][n]{\step_{#1}}
\newcommand{\steps}{\step^*}
\newcommand{\diverge}{{\Uparrow}}
\newcommand{\term}[1][-]{{\Downarrow_{#1}}}

@@

%%% Permissions
\newcommand{\plainperm}[1]{\textsc{#1}}
%\newcommand{\rwlx}{\plainperm{rwlx}}
\newcommand{\rwx}{\plainperm{rwx}}
\newcommand{\rx}{\plainperm{rx}}
%\newcommand{\rwlxo}{\plainperm{rwlxo}}
%\newcommand{\rwlo}{\plainperm{rwlo}}
%\newcommand{\rwl}{\plainperm{rwl}}
%\newcommand{\rwxo}{\plainperm{rwxo}}
%\newcommand{\rwo}{\plainperm{rwo}}
\newcommand{\rw}{\plainperm{rw}}
%\newcommand{\rxo}{\plainperm{rxo}}
\newcommand{\readonly}{\plainperm{r}}
\newcommand{\ro}{\readonly}
\newcommand{\noperm}{\plainperm{0}}
%\newcommand{\nopermo}{\plainperm{0o}}
%\newcommand{\enter}{\plainperm{e}}
%\newcommand{\entero}{\plainperm{eo}}

@@

\newcommand{\encType}[1]{\plainfun{encodeType}{#1}}
\newcommand{\encPerm}[1]{\plainfun{encocePerm}{#1}}
\newcommand{\decPerm}[1]{\plainfun{decodePerm}{#1}}
\newcommand{\encInstr}[1]{\plainfun{encodeInstruction}{#1}}
\newcommand{\decInstr}[1]{\plainfun{decodeInstruction}{#1}}
\newcommand{\updPcAddr}[1]{\plainfun{updatePc}{#1}}
%\newcommand{\updPcPerm}[1]{\plainfun{updatePcPerm}{#1}}
\newcommand{\nonExec}[1]{\plainfun{nonExecutable}{#1}}
\newcommand{\exec}[1]{\plainfun{executable}{#1}}
%\newcommand{\opaquePerm}[1]{\plainfun{opaquePerm}{#1}}
\newcommand{\readAllowed}[1]{\plainfun{readAllowed}{#1}}
\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}{#1}}
%\newcommand{\writeLocalAllowed}[1]{\plainfun{writeLocalAllowed}{#1}}
%\newcommand{\nonLoc}[1]{\plainfun{nonLocal}{#1}}
\newcommand{\nonLinear}[1]{\plainfun{nonLinear}{#1}}
\newcommand{\linCons}[1]{\plainfun{linearityConstraint}{#1}}
%\newcommand{\isLoc}[1]{\plainfun{isLocal}{#1}}
\newcommand{\isLinear}[1]{\plainfun{isLinear}{#1}}
\newcommand{\sealAss}[1]{\plainfun{sealAssignment}{#1}}
\newcommand{\noRetStkReg}[1]{\plainfun{noRetStk_{reg}}{#1}}
\newcommand{\noRetStkMs}[1]{\plainfun{noRetStk_{ms}}{#1}}
\newcommand{\withinBounds}[1]{\plainfun{withinBounds}{#1}}
\newcommand{\nonZero}[1]{\plainfun{nonZero}{#1}}
\newcommand{\callCond}[1]{\plainfun{callCondition}{#1}}

@@

% \dominique{7-9-2017: Should we keep local/global and linear/non-linear?  I was thinking just the latter would suffice...}
% \lau{07-09-2017: The local/global mechanism still has some value as it now can be used for the purpose of arguments that cannot be stored by the receiver (if you are willing to ``pay'' the clearing cost). I do, however, not think that it is strictly necessary here to make things work.}
% \lau{12-09-2017: I will remove the local capabilities today for the reasons we have discussed (they add to the complexity and they seem to be of little to no use without a clear memory instruction).}
Source language domains:
\[
  \begin{array}{rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
    &\Word & \defeq & \ints \uplus \SealableCaps\\
    \permbnf \in& \Perm & \defbnf & \dots \\
    &\linbnf & \defbnf & \linear \mid \normal \\
    &\aendbnf & \in & \Addr \uplus \{\infty \} \\
    &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
    &\scbnf &\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\linbnf,\sealbasebnf,\sealendbnf,\sigma}\\
    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd(\sigma,(\permbnf,\basebnf,\aendbnf,\addrbnf)) \mid \retptrc(\sigma,(\permbnf,\linbnf,\basebnf,\aendbnf,\addrbnf))}\\ 
    &\rbnf & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
    &\Reg & \defeq & \RegName \fun \Word\\
    &\Mem & \defeq & \Addr \fun \Word \\
    &\MemSeg & \defeq & \Addr \parfun \Word \\
    {\sourcecolor \sstkframe \in} & {\sourcecolor \StkFrame} & {\sourcecolor \defeq} & {\sourcecolor \Caps \times \MemSeg \times \Addr}\\
    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus (\{\halted\} \times \Mem)
  \end{array}
\]
\lau{06-09-2017: Should seals also have the option to be linear?}
\lau{13-09-2017: For now I have added linearity to the seals.}
\dominique{27-9-2017: I think we should have versions of $\retptrd$ and $\retptrc$ for non-top stack frames and define that invoking them with the current stack pointer will fail.}
\lau{28-09-2017: Why would you like this? If we want to add this, we will have to be careful, because an old code-return-pointer may fit perfectly well with our new data-return-pointer, so in this case it should not fail. We added the seal to the return pointers, so we had some way of figuring out whether two return pointers would actually fit together.}
\dominique{27-9-2017: Likewise, I think we need to have a source-language representation for a pointer into one of the non-top stack frames, as we may need to be able to back-translate adversary code that reads/writes in those.}
\lau{28-09-2017: At the moment, the only part of the stack that is stored in the stack frame is the private part. Specifically it is the part from the base to the current address of the stack pointer \emph{in $r_\stk$}. This means that if someone saves a stack pointer, so it is not included in the call, then they can still use it because that part of the stack was not encapsulated in a stack frame. Makes sense?}
The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
\begin{itemize}
\item $\linbnf$ defines domain $\Linear$
\item $\scbnf$ defines domain $\SealableCaps$
\item $\cbnf$ defines domain $\Caps$
\item $\rbnf$ defines the finite set $\RegName$.
\item $\Perm$ is defined as the set of permissions in Figure~\ref{fig:perm-hier}.
\end{itemize}

@@

\subsection{Permissions}
\lau{23-03-2017: I am not sure we need the opaque capabilities anymore.} % I am not sure why I did not think the opaque capabilities where necessary, but I have a feeling we talked about this in Leuven and came to the conclusion that they were necessary.
\begin{figure}[!h]
  \centering
  \begin{tikzpicture}[main node/.style={}]
    \node[main node] (rwx) {$\rwx$};
    \node[main node] (rx) [below right of=rwx] {$\rx$};

    \node[main node] (rw) [below left of=rwx,] {$\rw$};
    \node[main node] (r) [below right of=rw] {$\readonly$};
    \node[main node] (0) [below of=r] {$\noperm$};

@@

    (rwx) edge (rx)

    (rw) edge (rwx)
    (r) edge (rx);
  \end{tikzpicture}

@@

\begin{align*}
  \src{\Phi} & \; \src{\step \sem{\scall{r_1}{r_2}}(\Phi)} &  &\sourcecolor\left.\arraycolsep=0pt
                                                  \begin{array}[t]{l}
                                                    \text{if }\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr) \tand \\
                                                    \callCond{\Phi,r_1,r_2,\aaddr} \tand\\
                                                    \baddr \leq \aaddr \tand \aaddr + \calllen-1 \leq \eaddr \tand \\
                                                    \exec{\Phi(\pcreg)}
                                                  \end{array}\right.\\
  \Phi & \step \sem{\decInstr{\Phi.\mem(\aaddr)}}(\Phi) & &\left.\arraycolsep=0pt
                                                  \begin{array}[t]{l}
                                                    \text{if }\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr) \tand \\
                                                    \src{(\neg\callCond{\Phi,r_1,r_2,\aaddr} \tor}\\
                                                    \src{\eaddr < \aaddr + \calllen-1) \tand} \\
                                                    \baddr \leq \aaddr \leq \eaddr\tand\\
                                                    \exec{\Phi(\pcreg)}
                                                  \end{array}\right.\\
  \Phi& \step \failed & & \totherwise
\end{align*}

@@

@@

\noindent\textbf{jmp and jnz}\\
\begin{align*}
  \sem{\tjmp{r}}(\Phi) = &  
                     \begin{cases}
                       \Phi\updReg{r}{w}\updReg{\pcreg}{\Phi(r)} & w = \linCons{\Phi(r)}
                     \end{cases}
\end{align*}

@@

\noindent\textbf{cca}\\
\emph{Change Current Address}
\lau{15-09-2017: This is the old lea. I changed the name, so it actually reflects what the instruction does. We might also want to add a real lea at some point.}
\begin{align*}
  \sem{\tcca{r_1}{r_2}}(\Phi) = & 
                                  \begin{cases}
                                    \updPcAddr{\Phi\updReg{r_1}{c}} &  
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$}\\
                                      \quad\text{ and in either case $n \in \ints$ and} \\
                                      \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                      c = ((\perm,\lin),\baddr,\eaddr,\aaddr + n)
                                    \end{array}
                                    \\
                                    \updPcAddr{\Phi\updReg{r_1}{c}} &  
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$}\\
                                      \quad\text{ and in either case $n \in \ints$ and} \\
                                      \Phi(r_1) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma} \tand \\
                                      c = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma + n}
                                    \end{array}
                                    \\
                                    \src{\updPcAddr{\Phi\updReg{r_1}{c}}} &  
                                    \sourcecolor\left.
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \text{either $n = r_2$ or $n = \Phi(r_2)$}\\
                                      \quad\text{ and in either case $n \in \ints$ and} \\
                                      \Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
                                      c = \stkptr{\perm,\baddr,\eaddr,\aaddr + n}
                                    \end{array}\right.
                                    \\
                                    \failed & \totherwise
                                  \end{cases}
\end{align*}

@@

\noindent\textbf{seta2b}\\
\begin{align*}
  \sem{\tsetatob{r_1}{r_2}}(\Phi) = & 
                                \begin{cases}
                                  \updPcAddr{}\left(
                                    \arraycolsep=0pt
                                    \begin{array}{rl}
                                    \Phi&\updReg{r_2}{w_2}\\
                                        &\updReg{r_1}{c}
                                    \end{array} \right)
&
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\_) \tand \\
                                      w_2 = \linCons{\Phi(r_2)}\\
                                      c = ((\perm,\lin),\baddr,\eaddr,\baddr)
                                    \end{array} \\
                                  \updPcAddr{}\left(
                                    \arraycolsep=0pt
                                    \begin{array}{rl}
                                    \Phi&\updReg{r_2}{w_2}\\
                                        &\updReg{r_1}{c}
                                    \end{array} \right)
&
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \Phi(r_2) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\_} \tand \\
                                      w_2 = \linCons{\Phi(r_2)}\\
                                      c = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}
                                    \end{array} \\
                                    \failed & \totherwise
                                \end{cases}
\end{align*}

@@

@@

@@

\section{Logical Relation}
\lau{19-10-2017: Rest of source configuration needs to come from somewhere in $\lre$.}
\lau{19-10-2017: Maybe linearity should be taken care of in just one place (instead of in $\lrr$ relation and mem staisfaction). That is, add a predicate that checks whether the configuration satisfies linearity (such a condition may be to syntactic to work?).}
In the following definitions, \src{blue} is used to indicate values related to the source machine. This is unlike previous definitions, where \src{blue} is used  indicate source language specific parts of definitions.
\[
  \lre(W) = \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
    \begin{array}{l}
      \forall n' \leq n, \src{k_{d,S}}, \src{k_{c,S}}, k_{d,T}, k_{c,T}, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
      \quad \forall W_\lrk, W_\lrr, W_\lrm \ldotp \\
      \qquad\npair[n']{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \in \lrp(W) \wedge \\
      \qquad\npair[n']{\stpair[.]{k_{d,S},k_{c,S}}{k_{d,T},k_{c,T}}} \in \lrk(W_\lrk) \wedge \\
      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
      \qquad\Rightarrow\\
      \qquad\npair[n']{\left(\array{l}\src{(\ms_S,\reg_S\updReg{\pcreg,\rdata,\rretc,\rretd}{v_{c,S},v_{d,S},k_{c,S},k_{d,S}},\stk, \ms_\stk)},\\
                                          (\ms_T,\reg_T\updReg{\pcreg,\rdata,\rretc,\rretd}{v_{c,T},v_{d
,T},k_{c,T},k_{d,T}})\endarray\right)}
      \in \lro(W\uplus W_\lrk \uplus W_\lrr \uplus W_\lrm)
    \end{array}
    \right\}
\]
\lau{24-10-2017: Dominique, I do not see how the splitting of worlds is going to work out. I can see how it will enforce linearity, but is you give a "pure" region (the kind of region that is only allowed to go to one part of a split) to the $\lrr$-relation, then the memory satisfation does not have it, and it cannot find a memory segment for this part. In other words, the memory satisfaction will not be able to find a concrete memory segment for segments governed by a linear capability in the register file. Maybe the memory satisfaction should have all of the parts, so it can represent them (but also limit what linear capabilities can be used).}

\[
  \lro(W) = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
    \begin{array}{l}
      \forall i \leq n \ldotp \\
      \quad \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \term[i] \Rightarrow (\ms_T,\reg_T) \term\\
      \quad \tand\\
      \quad (\ms_T,\reg_T) \term[i] \Rightarrow \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \term
    \end{array}
\right\}
\]
